Gaussian Processes Autoregressive Models for the Disturbance Storm Time Index

### 2016-04-25 swsc.tex
Updates from ShareLaTeX

\begin{eqnarray}
    \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
    \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \\
    \Sigma_* & = & \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*}
\end{eqnarray}

@@

\begin{figure}[ht]
   \centering
   \subfigure{\includegraphics[width=\columnwidth]{Compare_RMSE.png}}
   \vspace{12pt}
   \subfigure{\includegraphics[width=\columnwidth]{Compare_CC.png}}
   \caption{\small  Performance Comparison; left: Averaged RMSE right: Averaged Cross correlation coefficient} 
   \label{fig:rmsecc}
   \end{figure}

@@

\begin{figure}
   \centering
   \subfigure{\includegraphics[width=\columnwidth]{Compare_deltaDst.png}}
   \vspace{12pt}
   \subfigure{\includegraphics[width=\columnwidth]{Compare_timingerr.png}}
   \caption{\small  Performance Comparison; left: Averaged $\Delta Dst_{min}$ right: Averaged absolute timing error} 
   \label{fig:deltaDst}
   \end{figure}

@@

For the purpose of generating the cumulative probability plots of the model errors, we must obtain hourly predictions for all the storm events using the \emph{NARMAX} and \emph{TL} models. In the case of the \emph{NM} model, the formula outlined in \citet{balikhin:narmax} is used to generate hourly predictions. For the \emph{TL} model we use real time predictions listed on their website corresponding to all the storm events. The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $D_{st}$ predictions for each hour.

@@

@@

\section{Results} \label{sec:res}
%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages

Figures \ref{fig:rmsecc} and \ref{fig:deltaDst} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}. Two central themes come out of these charts.

@@

@@

We conclude this paper with three key conclusions/insights learned from our activities so far.
   \begin{enumerate}
      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $D_{st}$ index. This is due to its excellent performance in terms of errors and correlations. 
      \item \emph{Gaussian Process} AR and ARX models give encouraring benefits in OSA prediction even when compared to the \emph{Persistence} model. Leveraging the strengths of the Bayesian approach, they are able to learn robust predictors from data.
      \item Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices. Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
   \end{enumerate}

>>>>

\begin{eqnarray}
    \mathbf{f_*}|\mathbf{X},\mathbf{y},\mathbf{X_*} & \sim & \mathcal{N}(\mathbf{\bar{f}_*}, \Sigma_*)  \label{eq:posterior} \\
    \mathbf{\bar{f}_*} & = & \mathbf{K}^T_{*} [\mathbf{K} + \sigma^{2} \mathbf{I}]^{-1} \mathbf{y} \label{eq:posteriormean} \\
    \Sigma_* & = & \mathbf{K}_{**} - \mathbf{K}^T_{*} \left(\mathbf{K} + \sigma^{2} \mathbf{I}\right)^{-1} \mathbf{K}_{*} \label{eq:posteriorcov}
\end{eqnarray}

The distribution of $\mathbf{f_*}| \mathbf{X},\mathbf{y},\mathbf{X_*}$ is known in Bayesian analysis research as the \emph{Posterior Predictive Distribution}. This illustrates a key difference between \emph{Gaussian Processes} and other regression models such as \emph{Neural Networks}, \emph{Linear Models} and \emph{Support Vector Machines}, that a \emph{Gaussian Process} model does not generate point predictions for new data but outputs a predictive distribution for the quantity sought, this allows us to construct error bars on the predictions. This property of Bayesian models such as \emph{Gaussian Processes} makes them suitable for Space Weather forecasting applications.

@@

\begin{figure}
   \centering
   \includegraphics[width=\textwidth]{Compare_RMSE.png}
      \caption{Averaged RMSE}
         \label{fig:rmse}
   \end{figure}

@@

@@

For the purpose of generating the cumulative probability plots of the model errors, we must obtain hourly predictions for all the storm events using the \emph{NARMAX} and \emph{TL} models. In the case of the \emph{NM} model, the formula outlined in \citet{balikhin:narmax} is used to generate hourly predictions. For the \emph{TL} model we use real time predictions listed on their website (\url{http://lasp.colorado.edu/home/spaceweather/}) corresponding to all the storm events. The real time predictions are at a frequency of ten minutes, hourly predictions are generated by averaging the $D_{st}$ predictions for each hour.

@@

\begin{figure}
   \centering
   \includegraphics[width=\textwidth]{Compare_CC.png}
      \caption{Averaged Cross correlation coefficient}
         \label{fig:cc}
   \end{figure}

@@

\begin{figure}
   \centering
   \includegraphics[width=\textwidth]{Compare_deltaDst.png}
      \caption{Averaged $\Delta Dst_{min}$}
         \label{fig:deltaDst}
   \end{figure}

\section{Results} \label{sec:res}
%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages
\begin{figure}
   \centering
   \includegraphics[width=\textwidth]{Compare_timingerr.png}
      \caption{Averaged absolute timing error}
         \label{fig:timingErr}
   \end{figure}

Figures \ref{fig:rmse}, \ref{fig:cc}, \ref{fig:deltaDst} and \ref{fig:timingErr} compare the performance of \emph{GP-AR}, \emph{GP-ARX} and the \emph{Persistence} model with the existing results of \citet{Ji2012}. Two central themes come out of these charts.

@@

We conclude this paper with three key conclusions/insights learned from our research so far.
   \begin{enumerate}
      \item \emph{Persistence} model must be central in the model evaluation process in the context of \emph{One Step Ahead} prediction of the $D_{st}$ index. This is due to its excellent performance in terms of errors and correlations. 
      \item \emph{Gaussian Process} AR and ARX models give encouraging benefits in OSA prediction even when compared to the \emph{Persistence} model. Leveraging the strengths of the Bayesian approach, they are able to learn robust predictors from data.

@@

      \item Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices. Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.
   \end{enumerate}
